---
title: Exploratory Data Analysis
format: typst
---

```{r}
library(tidyverse)
sov_clean <- read_csv("data/g24_sov_by_g24_svprec.csv")
```

### Question 1

What is the distribution of total votes ('TOTVOTE') across precincts?

### Answer 1

```{r}
sov_clean |>
    ggplot(aes(TOTVOTE)) +
    geom_histogram(bins = 40) +
    labs(title = "Distribution of Total Votes per Precinct")
```

Total votes are highly right-skewed. Most precincts have low-to-moderate turnout (under 500 votes), while a few aggregated units contain very large totals.

### Question 2

How does total turnout vary across congressional districts?

### Answer 2

```{r}
sov_clean |> 
  group_by(CDDIST) |> 
  summarize(total_votes = sum(TOTVOTE, na.rm = TRUE)) |> 
  ggplot(aes(CDDIST, total_votes)) +
  geom_col() +
  labs(title = "Total Votes by Congressional District")
```

Turnout varies widely across districts, reflecting population differences. Larger urban districts show higher total votes.

### Question 3

How do Democratic and Republican vote totals compare across precincts?

### Answer 3

```{r}
sov_clean |> 
  ggplot(aes(DEMVOTE, REPVOTE)) +
  geom_point(alpha = 0.4) +
  labs(title = "DEM vs REP Vote Totals per Precinct")
```

Democratic vote totals tend to exceed Republican vote totals in most precincts, though there are precincts where Republicans lead.